---
title: "Big Data Project 1"
author: "James Cevallos and Yash Huilgol"
date: "2/16/2021"
output: pdf_document
---

# Big Data Project One 

## Data Import
```{r data import, include=TRUE}
#import all csv files 
library(readr)
setwd("~/Desktop/HeritageHealth")
sampleTable <- read_csv("SampleEntry.csv")
setwd("~/Desktop/HeritageHealth/HHP_release3")
#used for prediction 
claims <- read_csv("Claims.csv") 
drugCount <- read_csv("DrugCount.csv")
labCount <- read_csv("LabCount.csv")
members <- read_csv("Members.csv")
#training data
daysInHospital_Y2 <- read_csv("DaysInHospital_Y2.csv")
#testing data
daysInHospital_Y3 <- read_csv("DaysInHospital_Y3.csv")
#descriptor tables not super helpful
lookUpPrimaryConditionGroup <- read_csv("Lookup PrimaryConditionGroup.csv")
lookUpProcedureGroup <- read_csv("Lookup ProcedureGroup.csv")
#not sure what this table is used for
target <- read_csv("Target.csv")
```

The goal of this project is to create an algorithm that will predict how many days a patient will spend in a hospital. 

To start out the project, we first calculated summary statistics and created a histogram to visualize the number of members who spent particular days in the hospital for Year 2. 
```{r}
library(dplyr)
library(ggplot2)

#summary statistics, not included in final report
count_data <- count(daysInHospital_Y2, DaysInHospital)
summary(daysInHospital_Y2$DaysInHospital)

#plot including 0 hospital stays, not included in final report 
temp_plot <- ggplot(daysInHospital_Y2, aes(DaysInHospital)) + geom_histogram(binwidth = .5, bins = 15)

#plot excluding 0 hospital stays, included in final report 
hospital_data_at_least_one_day <- daysInHospital_Y2 %>% filter(DaysInHospital > 0)
a <- ggplot(hospital_data_at_least_one_day , aes(DaysInHospital)) + geom_histogram(binwidth = .5, bins = 15) + scale_x_continuous(breaks= seq(0,15,1)) + labs(title = "Count of Total Number of Members by Days in Hospital") + xlab("Days in Hospital") + ylab("Number of Members") + theme(plot.title = element_text(hjust = 0.5))
a

# code to save plot to desktop
#ggsave(filename = "figure_1.jpeg", plot= a, device = "jpeg", path = "~/Desktop/")


```
In order to predict the number of days a patient will stay in the hospital, we decided that we'd like to capture data from the following five variables: 

1) The total number of claims during year 1 by member 
2) The total number of labs during year 1 by member 
3) The total number of prescriptions during year 1 by member 
4) The age at first claim for each member 
5) The total number of days in the hospital during year 2 by member 

To capture the data, we created a table with all this information by filtering and merging data from five tables: claims, drugCount, labCount, members, and daysInHospital_Y2

```{r}
#import dplyr library
library(dplyr)

#create a table with all the unique claims a member made for year one
year_1_claims <- claims %>% filter(Year == "Y1") %>% select(MemberID) %>% count(MemberID) %>% rename(number_of_claims=n)

#create a table of all the prescriptions a member has had filled for year one (Note all 7+ <- 7)
by_memberdrugs <- drugCount %>% filter(Year == "Y1") 
by_memberdrugs$DrugCount[by_memberdrugs$DrugCount == "7+"] <- "7"
by_memberdrugs$DrugCount <- as.numeric(by_memberdrugs$DrugCount)
by_memberdrugs <- group_by(by_memberdrugs, MemberID) %>% summarise(total_prescriptions_filled = sum(DrugCount))
year_1_drugs <- by_memberdrugs

#create a table of all the labs a member has completed for year one (Note all 10+ <- 10)
by_memberlabs <- labCount %>% filter(Year == "Y1")
by_memberlabs$LabCount[by_memberlabs$LabCount == "10+"] <- 10
by_memberlabs$LabCount <- as.numeric(by_memberlabs$LabCount)
by_memberlabs <- group_by(by_memberlabs, MemberID) %>% summarise(total_labs = sum(LabCount))
year_1_labs <- by_memberlabs

#merge tables together BEFORE CONTINUING MAKE SURE TO FILL IN NAs

temp_1 <- full_join(year_1_claims,year_1_drugs, by = "MemberID")
temp_1[is.na(temp_1)] <- 0
temp_1$total_prescriptions_filled <- as.integer(temp_1$total_prescriptions_filled)

temp_2 <- full_join(temp_1, year_1_labs, by = "MemberID")
temp_2[is.na(temp_2)] <- 0

temp_3 <- full_join(temp_2, members, by = "MemberID")
temp_3 <- na.omit(temp_3)

final_table <- full_join(temp_3, daysInHospital_Y2, by = "MemberID")
final_table <- select(final_table, -ClaimsTruncated)
final_table <- na.omit(final_table) #not sure if this is the best thing to do
```

Next, I plotted the data to visually inspect it in order to build our model 

```{r}

#make ints
final_table$number_of_claims <- as.integer(final_table$number_of_claims)
final_table$DaysInHospital <- as.integer(final_table$DaysInHospital)

ggplot(data=final_table, mapping = aes(x=total_prescriptions_filled, y=DaysInHospital)) + geom_point(size = 1, shape = 1) + xlim(0, 50) + ylim(0,20) 




```
